Evolution of \\ ice sheet geometry \\ using Stokes dynamics

### 2021-06-05 talk/talk.tex
Change sign of Phi.

\begin{block}{definition: ice dynamics operator}
$\Phi : s \mapsto \bus \cdot \bn_s$

@@

\begin{itemize}
\item to compute $\Phi(s)$: build $\Lambda_s$, solve weak-form Stokes problem
    $$F_{\Lambda_s}(\bu,p)[\bv,q] = \int_{\Lambda_s} 2 \nu_\eps D\bu : D\bv - p \Div\bv - (\Div\bu) q - \rhoi \bg \cdot \bv\,d\bx = 0$$

    \begin{itemize}
    \item well-posed over $W_0^{1,\pp}(\Lambda_s)^3 \times L^\qq(\Lambda_s)$ (Jouvet \& Rappaz \cite{JouvetRappaz2011})
    \item optimal solver exists (Isaac et al \cite{IsaacStadlerGhattas2015})
    \item extract trace $\bus$, then $\Phi(s) = \bus \cdot \bn_s$, extend by zero to all of $\Omega$
    \end{itemize}
\end{itemize}

@@

\begin{itemize}
\item using $\Phi$ gives a cleaner NCP over $\Omega$:
\begin{align*}
s - b &\ge 0 \\
- \Phi(s) - a &\ge 0 \\
(s - b) (- \Phi(s) - a) &= 0
\end{align*}

@@

\begin{block}{SIGP weak form}
find admissible $s \in \mathcal{K} = \{r \ge b\} \subset W^{1,\qq}(\Omega)$ so that
    $$\hspace{10mm} F(s)[r - s] \ge \ip{a}{r-s}$$
for all $r \in \mathcal{K}$, where $F(s)[w] = - \int_\Omega \Phi(s)\, w \,dx dy$
\end{block}

\item well-posedness is open
    \begin{itemize}
    \item existence holds for SIA version (Jouvet \& Bueler \cite{JouvetBueler2012})
    \end{itemize}

@@

\begin{block}{numerical goal}
Given triangulation $\mathcal{T}$ of $\Omega$ and $P_1$ space $\mathcal{V}^h \subset W^{1,\qq}(\Omega)$.  Use the numerical solution of the Stokes problem $F_{\Lambda_{s^h}}(\bu^h,p^h)=0$ to evaluate $\Phi(s^h)=\bu|_{s^h}\cdot\bn_{s^h}$.  Solve VI for $s^h \in \mathcal{K}^h$:

@@

\begin{itemize}
\item $F(s)[w] = - \int_\Omega \Phi(s)\, w$ is basically dual of $\Phi(s)=\bus\cdot\bn_s$
\item $\Phi : \mathcal{K} \subset W^{1,\qq}(\Omega) \to W^{-1,\pp}(\Omega)$
\item the Stokes stress balance is nonlocal so $\Phi$ is ``dense''
    \begin{itemize}
    \item some kind of integral operator (\emph{convolve with Stokeslets})
    \item $\Phi$ is concentrated near the diagonal, i.e.~short range
    \end{itemize}
\item yet it is a differential operator in the SIA case:
\small
    $$\Phi_{\text{SIA}}(s) = \frac{\gamma}{\qq} (s-b)^{\qq} |\grad_2 s|^{\qq} + \grad_2 \cdot\left[\frac{\gamma}{\qq+1} (s-b)^{\qq+1} |\grad_2 s|^{\qq-2} \grad_2 s\right]$$
\normalsize
\end{itemize}
\end{frame}

>>>>

\begin{block}{definition: ice dynamics operator}
$\Phi : s \mapsto - \bus \cdot \bn_s$

@@

\begin{itemize}
\item to compute $\Phi(s)$: build $\Lambda_s$, solve weak-form Stokes problem
    $$F_{\Lambda_s}(\bu,p)[\bv,q] = \int_{\Lambda_s} 2 \nu_\eps D\bu : D\bv - p \Div\bv - (\Div\bu) q - \rhoi \bg \cdot \bv\,d\bx = 0,$$
extract trace $\bus$, then extend $\Phi(s) = -\bus \cdot \bn_s$ by zero to $\Omega$
\item regarding $F_{\Lambda_s}(\bu,p) = 0$:
    \begin{itemize}
    \item well-posed over $W_0^{1,\pp}(\Lambda_s)^3 \times L^\qq(\Lambda_s)$ (Jouvet \& Rappaz \cite{JouvetRappaz2011})
    \item optimal solver exists (Isaac et al \cite{IsaacStadlerGhattas2015})
    \end{itemize}
\end{itemize}

@@

\begin{itemize}
\item using $\Phi$ gives a cleaner NCP over $\Omega$:
\begin{align*}
s - b &\ge 0 \\
\Phi(s) - a &\ge 0 \\
(s - b) (\Phi(s) - a) &= 0
\end{align*}

@@

\begin{block}{SIGP weak form}
find admissible $s \in \mathcal{K} = \{r \ge b\} \subset W^{1,\qq}(\Omega)$ so that
    $$\hspace{10mm} F(s)[r - s] \ge \ip{a}{r-s}$$
for all $r \in \mathcal{K}$, where $F(s)[w] = \int_\Omega \Phi(s)\, w \,dx dy$
\end{block}

\item well-posedness is completely open
    \begin{itemize}
    \item but existence holds for SIA version (Jouvet \& Bueler \cite{JouvetBueler2012})
    \end{itemize}

@@

\begin{block}{numerical goal}
Given triangulation $\mathcal{T}$ of $\Omega$ and $P_1$ space $\mathcal{V}^h \subset W^{1,\qq}(\Omega)$.  Use the numerical solution of the Stokes problem $F_{\Lambda_{s^h}}(\bu^h,p^h)=0$ to evaluate $\Phi(s^h)=-\bu|_{s^h}\cdot\bn_{s^h}$.  Solve VI for $s^h \in \mathcal{K}^h$:

@@

\begin{itemize}
\item $F(s)[w] = \int_\Omega \Phi(s)\, w$ is dual of $\Phi(s)=-\bus\cdot\bn_s$
\item $\Phi : \mathcal{K} \subset W^{1,\qq}(\Omega) \to W^{-1,\pp}(\Omega)$
\item the Stokes stress balance is nonlocal so $\Phi$ is ``dense''
    \begin{itemize}
    \item some kind of integral operator (\emph{convolve with bdry Stokeslets})
    \item $\Phi$ is concentrated near the diagonal, i.e.~short range?
    \end{itemize}
\item $\Phi \approx \Phi_{\text{SIA}}$ for shallow glaciers, a differential operator:
\small
    $$\Phi_{\text{SIA}}(s) = - \frac{\gamma}{\qq} (s-b)^{\qq} |\grad_2 s|^{\qq} - \grad_2 \cdot\left[\frac{\gamma}{\qq+1} (s-b)^{\qq+1} |\grad_2 s|^{\qq-2} \grad_2 s\right]$$
\normalsize
\end{itemize}
\end{frame}
